\section{To do thingies}\label{sec:ch2label}
\begin{itemize}
  \item Think about how to define real-time.
  \item Create an intuition for different parts of the equation
\end{itemize}

\subsection{Intuition for the frequency dependent damping term $2\sigma_1\partial_t\partial_x^2u$}
Take the frequency independent damping term $-2\sigma_0\partial_tu$. The more positive the velocity $\partial_tu$ is, i.e., the string is moving upwards the more this term applies a negative, or downwards force (/effect) on the string. Vice versa, a more negative velocity will make this term apply a more positive force on the string. As for the frequency dependent damping term, apart from the obvious $\sigma_1$, the effect of the term increases with an increase of $\partial_t\partial_x^2u$ which describes the rate of change of the curvature of the string.

Let's first talk about positive and negative curvature, i.e., when $\partial_x^2u > 0$ or $\partial_x^2u < 0$. Counterintuitively, in the positive case, the curve points downwards. Think about the function $f(x) = x^2$. It has a positive curvature (at any point), but has a minimum. We can prove this by taking $x=0$ and setting grid spacing $h=1$.
\begin{equation}
  \begin{aligned}
  \delta_{xx}f(x) &= \frac{1}{h^2} \left(f(-1)-2f(0)+f(1)\right), \\
  &= \frac{1}{1^2} \left((-1)^2-2\cdot0^2+1^2\right),\\
  &= \left(1-0+1\right) = 2.
  \end{aligned}
\end{equation}
In other words, the second derivative of the function $f(x)=x^2$ around $x=0$ is positive.

As our term does not only include a second-order spatial derivative but also a first-order time derivative, we are now talking about a positive or negative \textit{rate of change} of the curvature, i.e., when $\partial_t\partial_x^2u>0$ or $\partial_t\partial_x^2u<0$. A positive rate of change of curvature means that the string either has a positive curvature and is getting more positive, i.e., the string gets more curved over time, or that the string has a negative curvature and is getting less negative, i.e., the string gets less curved or 'loosens up' over time.  In the same way, a negative rate of change of curvature means that the string either has a negative curvature and is getting more negative, or that the string has a positive curvature and is getting less positive. 

Let's see some examples. Take the same function described before, but now $f$ changes over time, fx. $f(x, t)=tx^2$. When $t$ increases over time, the curvature gets bigger. Repeating what we did above with $x=0$ and grid spacing $h = 1$, but now with $t=2$ and step size $k=1$, but now with a backwards time derivative we get:
  \begin{alignat*}{3}
    \delta_{t-}\delta_{xx}f(x,t) &= \frac{1}{kh^2}\bigg(&&f(-1, 2) - 2f(0, 2) + f(1, 2) \\
    & &&- \Big(f(-1, 1) - 2f(0, 1) + f(1, 1)\Big)\bigg),\\
    & = \frac{1}{1\cdot 1^2}\bigg(&&2\cdot(-1)^2-2\cdot2\cdot(0)^2+2\cdot1^2\\
    & &&-\Big(1\cdot(-1)^2-2\cdot1\cdot(0)+1\cdot(1^2)\Big)\Bigg),\\
    &=2+2-&&(1+1)=2.
  \end{alignat*}
So the rate of change of the curvature is positive, i.e., the already positively curved function $x^2$ gets more curved over time.

If the curvature around a point along a string gets more positive (or less negative) over time, the force applied to that point will be positive. Vice versa, if the curvature around a point along a string gets more negative (or less positive) over time, the force applied will be negative.
 
The fact the frequency dependent term to be added rather than subtracted to the FDS, is caused by the fact that a positive curvature implies a negative position in the string (think of the function $x^2$ which has a positive curvature, but the function `points' downwards). This translated to the force/effect this term has on the scheme means...
