\chapter*{Resonators}
<<<<<<< HEAD
Though the physical models described in the previous part are also considered resonators, they are \textit{ideal} cases. In other words, you would not be able to find these ``in the wild'' as they do not includes effects such as losses or frequency dispersion. 

This part presents the different resonators used over the course of the project and is structured as follows: Chapter \ref{ch:stiffString} introduces the stiff string, a model which has been used a lot over the course of this project, Chapter \ref{ch:brass} talks about brass instruments, or more generally, 1D systems of varying geometry along their spatial dimension. Finally, Chapter \ref{ch:2Dsyst} will introduces 2D systems which, in this project, have been used to simulate (simplified) instrument bodies.
=======
Although the physical models described in the previous part -- the simple mass-spring system and the 1D wave equation -- are also considered resonators, they are \textit{ideal} cases. In other words, these can not be found in the real world as effects such as losses or frequency dispersion are not included. 

This part presents the different resonators used over the course of the project that better include these non-ideal physical processes and is structured as follows: Chapter \ref{ch:stiffString} introduces the stiff string, an extension of the 1D wave equation, % and is the single most-used model in this project. 
Chapter \ref{ch:brass} introduces acoustic tubes, used to model brass instruments, and finally, Chapter \ref{ch:2Dsyst} introduces 2D systems which, in this project, have been used to simulate (simplified) instrument bodies. The analysis techniques introduced in the previous part will be applied to all models and described in detail. 
>>>>>>> master
